---
title: "Analysis, Visualization and Mapping in R"
subtitle: "Exercise 3: Descriptive Statistics"
output: 
  html_document:
    code_folding: hide
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include=FALSE, warning=FALSE, message=FALSE, eval=FALSE)
```

1. Create a new Rmarkdown document in Rstudio. Load the 'tidyverse' and 'knitr' packages. Install and load the 'devtools' and then the 'CEPESP-R' package. 
```{r}
library(tidyverse)
library(knitr)
#install.packages("devtools")
library(devtools)
#devtools::install_github("Cepesp-Fgv/cepesp-r")
library(cepespR)
```

2. Let's use the CEPESP-R API to download the voting data from the 2016 municipal elections. See [here](https://github.com/Cepesp-Fgv/cepesp-r) and the code below.

```{r, echo=TRUE}
data <- cepespdata(year=2016, position="Prefeito", regional_aggregation="Municipality",political_aggregation="Candidate")
```

3. On average, how many people voted in a municipality in the first round of the 2016 election?

```{r}
data %>% filter(NUM_TURNO==1) %>% 
  group_by(COD_MUN_IBGE) %>% 
  summarize(total_votes=sum(QTDE_VOTOS)) %>% 
  summarize(avg_votes=mean(total_votes))
```

4. What percentage of the votes cast in each region (`NOME_MACRO`) of the country were won by PMDB prefeito candidates? 

```{r}
data %>% group_by(NOME_MACRO,SIGLA_PARTIDO) %>%
  summarize(total_votes=sum(QTDE_VOTOS)) %>%
  group_by(NOME_MACRO) %>%
  mutate(pct_votes=100*(total_votes/sum(total_votes))) %>%
  filter(SIGLA_PARTIDO=="PMDB") %>%
  kable(digits=1)
  
```

5. Now compare the previous result to the percentage of elected prefeitos (`DESC_SIT_TOT_TURNO`) who were from the PMDB in each region. 

```{r}
data %>% filter(DESC_SIT_TOT_TURNO=="ELEITO") %>%
  group_by(NOME_MACRO,SIGLA_PARTIDO) %>%
  count() %>%
  group_by(NOME_MACRO) %>%
  mutate(pct_votes=100*(n/sum(n))) %>%
  filter(SIGLA_PARTIDO=="PMDB") %>%
  kable(digits=1)
```

6. How do male candidates differ from female candidates (`DESCRICAO_SEXO`)? Compare averages on two variables: The number of votes (`QTDE_VOTOS`) and age (`IDADE_DATA_ELEICAO`). Report the averages in a simple table.

```{r}
vars <- c("QTDE_VOTOS","IDADE_DATA_ELEICAO")

data %>% filter(NUM_TURNO==1) %>%
  group_by(DESCRICAO_SEXO) %>%
  summarise_at(vars,mean) %>%
  kable(digits=1)
```

7. Conduct a t-test to assess if women prefeitos receive a statistically different quantity of votes to men. Report a simple table of the results. Hint: Remember the `tidy()` action after a messy statistical test.

```{r}
data %>% filter(NUM_TURNO==1) %>%
  t.test(QTDE_VOTOS ~ DESCRICAO_SEXO, data=.) %>%
  tidy() %>%
  kable()
```

8. Conduct a difference-in-means t-test between men and women of both number of votes and age at the same time. Report the p-values of each t-test

```{r}
data %>% filter(NUM_TURNO==1) %>%
  summarise_at(vars,funs(t.test(.[DESCRICAO_SEXO=="MASCULINO"],.[DESCRICAO_SEXO=="FEMININO"])$p.value)) %>%
  round(3) %>%
  kable()
```

9. Create a neat, well-labelled, table of the average age of elected prefeitos in each region (`NOME_MACRO`) and by gender (`DESCRICAO_SEXO`). The table should have regions in the rows and gender in the columns.
```{r}
data %>% filter(DESC_SIT_TOT_TURNO=="ELEITO") %>%
  group_by(NOME_MACRO,DESCRICAO_SEXO) %>%
  summarize(Average_Age=mean(IDADE_DATA_ELEICAO)) %>%
  spread(key="DESCRICAO_SEXO",value="Average_Age") %>%
  kable(caption="Average Age of Elected Prefeitos by Region and Gender", align=c("lcc"), row.names=FALSE, col.names=c("Region","Women","Men"))
```

10. Are married candidates (`DESCRICAO_ESTADO_CIVIL`) for prefeito more likely to be elected than unmarried (all groups except married) candidates? Provide a well-formatted, labelled, table.
```{r}
table(data$DESCRICAO_ESTADO_CIVIL)

data %>% group_by(DESCRICAO_ESTADO_CIVIL,DESC_SIT_TOT_TURNO) %>%
  count() %>%
  group_by(DESCRICAO_ESTADO_CIVIL) %>%
  mutate(Pct_Elected=round(100*(n/sum(n)),1)) %>%
  filter(DESC_SIT_TOT_TURNO=="ELEITO") %>%
  select(DESCRICAO_ESTADO_CIVIL,Pct_Elected) %>%
  kable(caption="Rate of Election by Marital Status",align="lc",col.names=c("Marital Status","% Elected"))

```

