---
title: "Analysis, Visualization and Mapping in R"
subtitle: "Exercise 2: Manipulating a Big Dataset"
output: 
  html_document:
    code_folding: hide
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include=TRUE, warning=FALSE, message=FALSE, eval=FALSE)
```

1. Create a new Rmarkdown file in Rstudio.
2. Load the 'tidyverse', 'knitr' and 'nycflights13' packages
```{r}
library(tidyverse)
library(knitr)
library(nycflights13)
```

3. Starting from the `flights` dataset, create a table of all the flights of airline carrier 'OO'.
```{r}
flights %>% filter(carrier=="OO") %>%
  kable()
```

4. What is the average departure delay across all flights?
```{r}
flights %>% summarize(avg_delay=mean(dep_delay,na.rm=TRUE)) %>%
  kable()
```

5. Calculate the speed of each flight and then report a table of just the carrier, flight and speed variables, and just the first 5 rows of the table.
```{r}
flights %>% mutate(speed=distance/air_time) %>% 
  select(carrier, flight, speed) %>%
  slice(1:5) %>%
  kable()
```

6. Calculate the average speed separately for each airline (carrier), and arrange the table from fastest to slowest.
```{r}
flights %>% mutate(speed=distance/air_time) %>% 
  group_by(carrier) %>%
  summarize(avg_speed=mean(speed,na.rm=TRUE)) %>%
  arrange(-avg_speed) %>%
  kable()
```

7. How do departure delays vary across months?
```{r}
flights %>% group_by(month) %>%
  summarize(avg_delay=mean(dep_delay,na.rm=TRUE)) %>%
  kable()
```

8. Make a table showing average departure delays by airline (row) and origin airport (column)
```{r}
flights %>% group_by(carrier,origin) %>%
  summarize(avg_delay=round(mean(dep_delay,na.rm=TRUE),2)) %>%
  spread(key="origin",value="avg_delay") %>%
  kable()
```

9. How many flights from JFK to San Francisco (SFO) landed after 11.45pm on time?

```{r}
flights %>% filter(origin=="JFK" & dest=="SFO" & arr_time>2345 & arr_delay<=0) %>%
  count()
```

10. How many distinct planes (each plane has a different `tailnum`) flew from Newark (EWR) to Phoenix (PHX) in August? Hint: use the verb `distinct`

```{r}
flights %>% filter(origin=="EWR" & dest=="PHX" & month==8) %>%
  distinct(tailnum) %>%
  count()
```

11. What percentage of flights that took off late arrived on time? Hint: Try making two separate calculations and then calculating the percentage.

```{r}
num_flights <- flights %>% filter(dep_delay>=0) %>%
  count()

late_flights <- flights %>% filter(dep_delay>=0 & arr_delay<=0) %>%
  count()

late_flights/num_flights*100
```
