---
title: "Analysis, Visualization and Mapping in R"
subtitle: "Exercise 3: Descriptive Statistics"
output: 
  html_document:
    code_folding: show
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include=TRUE, warning=FALSE, message=FALSE, eval=FALSE, echo=FALSE)
```

1. Create a new Rmarkdown document in Rstudio. Load the `tidyverse`, `zeligverse`, `knitr`, `stargazer` and `cepespR` packages. 
```{r, echo=TRUE}
library(tidyverse)
library(zeligverse)
library(knitr)
library(cepespR)
library(stargazer)
```

2. Let's use the CEPESP-R API to download the prefeito candidate voting data from the 2016 municipal elections. See [here](https://github.com/Cepesp-Fgv/cepesp-r) and the code below.

```{r, echo=TRUE}
data <- cepespdata(year=2016, position="Prefeito", regional_aggregation="Municipality",political_aggregation="Candidate")
```

3. Run a regression to assess if men get more votes than women (`DESCRICAO_SEXO`) in the first round of the election (`NUM_TURNO`), controlling for race (`DESCRICAO_COR_RACA`).

```{r}
data %>% filter(NUM_TURNO==1) %>% 
  zelig(QTDE_VOTOS ~ DESCRICAO_SEXO + DESCRICAO_COR_RACA, data=., model="ls")
```

4. Create a dummy variable (0/1) for whether each candidate was elected (`COD_SIT_TOT_TURNO` is `ELEITO`). Run the appropriate regression to assess if men are more likely to be elected than women, controlling for race (`DESCRICAO_COR_RACA`).

```{r}
data %>% filter(NUM_TURNO==1) %>% 
  mutate(Elected=ifelse(DESC_SIT_TOT_TURNO=="ELEITO",1,0)) %>%
  zelig(Elected ~ DESCRICAO_SEXO + DESCRICAO_COR_RACA, data=., model="logit")
```

5. Produce a neatly formatted table of the regression in question 4 using `stargazer`.

```{r, results='asis'}
reg_non <- data %>% filter(NUM_TURNO==1) %>% 
  mutate(DESCRICAO_SEXO=factor(DESCRICAO_SEXO),DESCRICAO_COR_RACA=factor(DESCRICAO_COR_RACA)) %>%
  mutate(Elected=ifelse(DESC_SIT_TOT_TURNO=="ELEITO",1,0)) %>%
  zelig(Elected ~ DESCRICAO_SEXO + DESCRICAO_COR_RACA, data=., model="logit") 

data_new <- data.frame(
  DESCRICAO_SEXO=c("MASCULINO","MASCULINO","FEMININO","FEMININO"),
  DESCRICAO_COR_RACA=c("PRETA","BRANCA","BRANCA","PRETA")
)

temp <- reg_non %>% setx(data_new) %>% sim() %>% zelig_qi_to_df() 


reg_non %>%
  from_zelig_model() %>%
  stargazer(type="html",title="The Effect of Gender on Probability of being Elected",digits=2)


reg_non %>% setx(DESCRICAO_SEXO="FEMININO",DESCRICAO_COR_RACA="PRETA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(fem_preta=mean(predicted_value))
 
reg_non %>% setx(DESCRICAO_SEXO="FEMININO",DESCRICAO_COR_RACA="BRANCA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(fem_branca=mean(predicted_value))

reg_non %>% setx(DESCRICAO_SEXO="MASCULINO",DESCRICAO_COR_RACA="PRETA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(masc_preta=mean(predicted_value))

reg_non %>% setx(DESCRICAO_SEXO="MASCULINO",DESCRICAO_COR_RACA="BRANCA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(masc_branca=mean(predicted_value))

```

6. Does being both black *and* female make it even harder to get elected? Use an interaction term in a regression to assess this hypothesis. Show the formatted output table.

```{r, results='asis'}
reg <- data %>% filter(NUM_TURNO==1) %>% 
  mutate(DESCRICAO_SEXO=factor(DESCRICAO_SEXO),DESCRICAO_COR_RACA=factor(DESCRICAO_COR_RACA)) %>%
  mutate(Elected=ifelse(DESC_SIT_TOT_TURNO=="ELEITO",1,0)) %>%
  zelig(Elected ~ DESCRICAO_SEXO*DESCRICAO_COR_RACA, data=., model="logit")

reg %>% from_zelig_model() %>%
  stargazer(type="html",title="The Effect of Gender on Probability of being Elected",digits=2)

reg %>% setx(DESCRICAO_SEXO="FEMININO",DESCRICAO_COR_RACA="PRETA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(fem_preta=mean(predicted_value))

reg %>% setx(DESCRICAO_SEXO="FEMININO",DESCRICAO_COR_RACA="BRANCA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(fem_branca=mean(predicted_value))

reg %>% setx(DESCRICAO_SEXO="MASCULINO",DESCRICAO_COR_RACA="PRETA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(masc_preta=mean(predicted_value))

reg %>% setx(DESCRICAO_SEXO="MASCULINO",DESCRICAO_COR_RACA="BRANCA") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(masc_branca=mean(predicted_value))

  
```

7. Use your regression in question 3 to predict how many votes on average an indigenous man would receive in an average election for prefeito. (Remember to make both explanatory variables factors before you run the regression).

```{r}
data %>% filter(NUM_TURNO==1) %>% 
  mutate(DESCRICAO_SEXO=factor(DESCRICAO_SEXO),DESCRICAO_COR_RACA=factor(DESCRICAO_COR_RACA)) %>%
  zelig(QTDE_VOTOS ~ DESCRICAO_SEXO + DESCRICAO_COR_RACA, data=., model="ls") %>%
  setx(DESCRICAO_SEXO="MASCULINO") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  summarize(avg_votes=mean(predicted_value))
```

8. Create a histogram (`geom_histogram`) of the full set of predictions from question 7. 

```{r}
data %>% filter(NUM_TURNO==1) %>% 
  mutate(DESCRICAO_SEXO=factor(DESCRICAO_SEXO),DESCRICAO_COR_RACA=factor(DESCRICAO_COR_RACA)) %>%
  zelig(QTDE_VOTOS ~ DESCRICAO_SEXO + DESCRICAO_COR_RACA, data=., model="ls") %>%
  setx(DESCRICAO_SEXO="MASCULINO") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  ggplot() +
  geom_histogram(aes(x=predicted_value)) +
  theme_classic()
```

9. Can you see anything unusual or inconsistent in the predicted values shown in your histogram in Question 8? Run the same code again but change the model to a `poisson` model that only predicts non-negative integer values. Which histogram makes more sense?

```{r}
data %>% filter(NUM_TURNO==1) %>% 
  mutate(DESCRICAO_SEXO=factor(DESCRICAO_SEXO),DESCRICAO_COR_RACA=factor(DESCRICAO_COR_RACA)) %>%
  zelig(QTDE_VOTOS ~ DESCRICAO_SEXO + DESCRICAO_COR_RACA, data=., model="poisson") %>%
  setx(DESCRICAO_SEXO="MASCULINO") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  ggplot() +
  geom_histogram(aes(x=predicted_value)) +
  theme_classic()
```

10. How does the number of votes received change if we change our indigenous man to an indigenous woman? (Hint: Use `setx1` after `setx` to specify a second set of x values. Use either a `ls` or `poisson` model.) Report the average of the predicted values for each gender.

```{r}
data %>% filter(NUM_TURNO==1) %>% 
  mutate(DESCRICAO_SEXO=factor(DESCRICAO_SEXO),DESCRICAO_COR_RACA=factor(DESCRICAO_COR_RACA)) %>%
  zelig(QTDE_VOTOS ~ DESCRICAO_SEXO + DESCRICAO_COR_RACA, data=., model="poisson") %>%
  setx(DESCRICAO_SEXO="MASCULINO") %>%
  setx1(DESCRICAO_SEXO="FEMININO") %>%
  sim() %>%
  zelig_qi_to_df() %>%
  group_by(DESCRICAO_SEXO) %>%
  summarize(avg_votes=mean(predicted_value)) %>%
  kable()
```
