---
title: "Analysis, Visualization and Mapping in R"
subtitle: "Exercise 1: Produce a quick ready-to-share analysis using cepespata"
output: 
  html_document:
    code_folding: hide
css: styles.css
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include=TRUE, warning=FALSE, message=FALSE, eval=FALSE)
```

1. Create a project in Rstudio.

2. Download a CSV from [cepesp.io](cepesp.io) to your project folder. Use the 'Eleicoes por Cargo' tab with the default options. The file should include the results of the 2014 Presidential election aggregated for the whole of Brazil.

3. Create a new Rmarkdown document file. 

4. Install and load the 'tidyverse' and 'knitr' packages
```{r}
install.packages(c("tidyverse","knitr"))
library(tidyverse)
library(knitr)
```

```{r, eval=TRUE, results='hide'}
library(tidyverse)
library(knitr)
```

5. Import the CSV to Rstudio.
```{r, eval=TRUE}
pres_2014_elec_results <- read_csv("TSE_PRESIDENTE_BR_CANDIDATO_2014.csv")
```

6. How many rows are there in the dataset?
```{r}
pres_2014_elec_results %>% count()
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output"> Output </button>  
<div id="Output" class="collapse"> 

```{r, eval=TRUE, echo=FALSE}
pres_2014_elec_results %>% count()
```

</div>

7. Note your dataset contains both the first and second rounds. How many candidates competed in the first round?
```{r}
pres_2014_elec_results %>% filter(NUM_TURNO==1) %>%
  count()
```
<button class="btn btn-primary" data-toggle="collapse" data-target="#Output2"> Output </button>  
<div id="Output2" class="collapse"> 

```{r, eval=TRUE, echo=FALSE}
pres_2014_elec_results %>% filter(NUM_TURNO==1) %>%
  count()
```
</div>

8. How many people voted in the first round of the election?

```{r}
voters_first <- pres_2014_elec_results %>%
  filter(NUM_TURNO==1) %>% 
  summarize(Total_Votes=sum(QTDE_VOTOS)) %>%
  pull(Total_Votes)
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output3"> Output </button>  
<div id="Output3" class="collapse"> 

```{r, eval=TRUE, echo=FALSE}
voters_first <- pres_2014_elec_results %>%
  filter(NUM_TURNO==1) %>% 
  summarize(Total_Votes=sum(QTDE_VOTOS)) %>%
  pull(Total_Votes)
voters_first
```
</div>

9. What was the percentage increase in the total number of people who voted between the first and second round? Use an in-line code expression.

```{r}
voters_second <- pres_2014_elec_results %>%
  filter(NUM_TURNO==2) %>% 
  summarize(Total_Votes=sum(QTDE_VOTOS)) %>%
  pull(Total_Votes)

pct_change <- 100*(voters_second - voters_first)/voters_first
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output4"> Output </button>  
<div id="Output4" class="collapse"> 
```{r, eval=TRUE, echo=FALSE}
voters_second <- pres_2014_elec_results %>%
  filter(NUM_TURNO==2) %>% 
  summarize(Total_Votes=sum(QTDE_VOTOS)) %>%
  pull(Total_Votes)

pct_change <- round(100*(voters_second - voters_first)/voters_first,1)
```

The percentage increase in the total number of people who voted between the first and second round was `r pct_change`%.
</div>

10. The data contains only the number of votes. Calculate the *vote share* of each candidate for each round as a new column in your data frame. (Remember to assign the result as an object we can use for the rest of the exercise).
```{r}
pres_2014_elec_results <- pres_2014_elec_results %>% 
  group_by(NUM_TURNO) %>% 
  mutate(Vote_Share=100*(QTDE_VOTOS/sum(QTDE_VOTOS)))
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output5"> Output </button>  
<div id="Output5" class="collapse"> 
```{r, eval=TRUE, echo=FALSE}
pres_2014_elec_results <- pres_2014_elec_results %>% 
  group_by(NUM_TURNO) %>% 
  mutate(Vote_Share=100*(QTDE_VOTOS/sum(QTDE_VOTOS)))

pres_2014_elec_results %>% select(NOME_CANDIDATO,Vote_Share)
```
</div>

11. Insert an 'in-line' sentence that explains which candidate won the election and what percentage of the vote they received. 
```{r}
winner <- pres_2014_elec_results %>% 
  filter(NUM_TURNO==2) %>%
  arrange(-Vote_Share) %>%
  slice(1) %>%
  select(SIGLA_PARTIDO, NOME_CANDIDATO, Vote_Share)

winner_name <- winner %>% pull(NOME_CANDIDATO)
winner_Votes <- winner %>% pull(Vote_Share)
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output6"> Output </button>  
<div id="Output6" class="collapse"> 
```{r, eval=TRUE, echo=FALSE}
winner <- pres_2014_elec_results %>% 
  filter(NUM_TURNO==2) %>%
  arrange(-Vote_Share) %>%
  slice(1) %>%
  select(SIGLA_PARTIDO, NOME_CANDIDATO, Vote_Share)

winner_name <- winner %>% pull(NOME_CANDIDATO)
winner_Votes <- winner %>% pull(Vote_Share) %>% round(1)
```

The winner was `r winner_name` with `r winner_Votes`% of the vote.
</div>

12. Produce a table which shows the results of the first round of the election, including vote share, but excluding boring columns like the year, candidate number etc.
```{r}
pres_2014_elec_results %>% 
  filter(NUM_TURNO==1) %>% 
  select(NOME_CANDIDATO, SIGLA_PARTIDO, Vote_Share) %>%
  kable()
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output7"> Output </button>  
<div id="Output7" class="collapse"> 
```{r, eval=TRUE, echo=FALSE}
pres_2014_elec_results %>% 
  filter(NUM_TURNO==1) %>% 
  select(NOME_CANDIDATO, SIGLA_PARTIDO, Vote_Share) %>% kable()
```
</div>

13. Produce a bar chart which shows the results of the second round of the election by vote share.

```{r}
pres_2014_elec_results %>% 
  filter(NUM_TURNO==2) %>% 
  ggplot() + 
  geom_col(aes(x=NOME_CANDIDATO,y=Vote_Share))
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output8"> Output </button>  
<div id="Output8" class="collapse"> 
```{r, eval=TRUE, echo=FALSE}
pres_2014_elec_results %>% 
  filter(NUM_TURNO==2) %>% 
  ggplot() + 
  geom_col(aes(x=NOME_CANDIDATO,y=Vote_Share))
```
</div>

14. Specify whether you want the processing code to show up in your final document.
```{r}
#Place this line in a chunk at the start of your code to make sure only the outputs appear
opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

15. Produce a PDF which combines all your outputs (text, table, chart).

16. Produce a web page which combines all your outputs (text, table, chart).

17. (If you have time...) Include a line chart which shows how the vote share of the two candidates that made the second round changed from the first to the second round. 
```{r}
runoff_cands <- pres_2014_elec_results %>% 
  filter(NUM_TURNO==2) %>%
  pull(NUMERO_CANDIDATO)

pres_2014_elec_results %>% filter(NUMERO_CANDIDATO %in% runoff_cands) %>%
  ggplot() + geom_line(aes(x=as.factor(NUM_TURNO),y=Vote_Share,group=NOME_CANDIDATO,color=NOME_CANDIDATO)) +
  theme_classic() + 
  xlab("Round") +
  ylab("Vote Share")
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Output9"> Output </button>  
<div id="Output9" class="collapse"> 
```{r, eval=TRUE, echo=FALSE}
runoff_cands <- pres_2014_elec_results %>% 
  filter(NUM_TURNO==2) %>%
  pull(NUMERO_CANDIDATO)

pres_2014_elec_results %>% filter(NUMERO_CANDIDATO %in% runoff_cands) %>%
  ggplot() + geom_line(aes(x=as.factor(NUM_TURNO),y=Vote_Share,group=NOME_CANDIDATO,color=NOME_CANDIDATO)) +
  theme_classic() + 
  xlab("Round") +
  ylab("Vote Share")
```
</div>