---
title: "Analysis, Visualization and Mapping in R"
subtitle: "Exercise 9: Shiny"
output: 
  html_document:
    code_folding: show
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include=TRUE, warning=FALSE, message=FALSE, eval=FALSE, echo=FALSE)
```

1. Create a new Shiny App in Rstudio. Load the `shiny`, `tidyverse` and `cepespR` packages.

```{r, echo=T}
library(shiny)
library(tidyverse)
library(cepespR)
```

2. Using the code below, download the electoral data file ONCE using `cepespR`, save the data as a ".csv" and then include a line opening the ".csv". (Remove the `cepespR` and `write.csv` code after).

```{r, echo=T}
data <- get_elections(year=2014, position="President",regional_aggregation="State",political_aggregation="Candidate") %>%
  mutate(NOME_PARTIDO=iconv(NOME_PARTIDO, "ASCII", "UTF-8", sub=""))

write.csv(data,"data.csv")
data <- read.csv("data.csv")
```

3. Before creating the Shiny app, create an example analysis first: Filter the election for the second round (`NUM_TURNO`) and the PSDB (`SIGLA_PARTIDO`). In a separate line of code, create a `ggplot` output that displays a column chart (`geom_col`) of the number of votes the party received in each state.

4. Now, let's put this analysis within a Shiny app. Create a User Interface with appropriate inputs to allow the user to select the round of the election (`NUM_TURNO`) and the Party (`SIGLA_PARTIDO`). 

5. Turn your code for filtering the dataset by round and party into a `reactive` function within the `server` function that responds to the two inputs in your UI.

6. Turn your ggplot2 graphic output into a `renderPlot` function within the `server` function so that it depends on the `reactive` function you created in the previous question.

7. Create a place for this `renderPlot` graphic to appear in your `ui` function using `plotOutput`. 

8. Test your Shiny App and see if it works!

9. Make the scale of your graph fixed (`ylim`) so it shows all the values but is easier to compare vote quantities when selecting between parties. (Hint: The maximum number of votes is for the PSDB in SP in the second round).

10. Create another `reactive` output that calculates the total percentage of the vote the selected party received in the selected round of the election. 

11. Create a text output (`renderText`) in the `server` function and a `textOutput` in the `ui` function to display the result of the calculation in Question 7.