---
title: "Analysis, Visualization and Mapping in R"
subtitle: "Exercise 1: Produce a quick ready-to-share analysis using cepespata"
output: 
  html_document:
    code_folding: hide
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include=TRUE, warning=FALSE, message=FALSE, eval=FALSE)
```

1. Create a new Rmarkdown file in Rstudio.
2. Load the 'tidyverse', 'knitr' and 'nycflights13' packages
```{r}
library(tidyverse)
library(knitr)
library(nycflights13)
```

3. Starting from the `flights` dataset, create a table of all the flights of airline carrier 'OO'.
```{r}
flights %>% filter(carrier=="OO")
```

4. Starting from the `flights` dataset, calculate the average departure delay across all flights.
```{r}
flights %>% summarize(avg_delay=mean(dep_delay,na.rm=TRUE))
```

5. Starting from the `flights` dataset, calculate the speed of each flight and then report a table of just the carrier, flight and speed variables, and just the first 5 rows of the table.
```{r}
flights %>% mutate(speed=distance/air_time) %>% 
  select(carrier, flight, speed) %>%
  slice(1:5)
```

6. Starting from the `flights` dataset, calculate the average speed separately for each airline (carrier), and arrange the table from fastest to slowest.
```{r}
flights %>% mutate(speed=distance/air_time) %>% 
  group_by(carrier) %>%
  summarize(avg_speed=mean(speed,na.rm=TRUE)) %>%
  arrange(-avg_speed)
```

7. Starting from the `flights` dataset, 
